% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{mathptmx}
\usepackage{float}
\usepackage{booktabs}
\usepackage[version=4]{mhchem}
% \usepackage{longtable}
\usepackage{caption}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
% \usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
% \usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{makecmds}

\newcommand{\COO}{\ce{CO2}}
\newcommand{\methane}{\ce{CH4}}
\newcommand{\degC}{^\circ \mathrm{C}}
\newcommand{\degF}{^\circ \mathrm{F}}
\newcommand{\water}{\mathrm{H_2O}}
\newcommand{\carb}{\ce{CO3^2-}}
\newcommand{\bicarb}{\ce{HCO3-}}
\newcommand{\carbonic}{\ce{H2CO3}}
\newcommand{\Hplus}{\ce{H+}}
\newcommand{\OH}{\ce{OH-}}
\newcommand{\silica}{\ce{SiO2}}
\newcommand{\calcite}{\ce{CaCO3}}
\newcommand{\Caplus}{\ce{Ca^2+}}
\newcommand{\silicate}{\ce{SiO3^2-}}
\newcommand{\CaSi}{\ce{CaSiO3}}
\newcommand{\pH}{p\ce{H}}
\newcommand{\permil}{\permille}

\newcommand{\NetLogo}{{\sffamily\scshape NetLogo}}
\newcommand{\NetLogoURL}{\url{https://ccl.northwestern.edu/netlogo/}}

\newcommand{\alintertext}[1]{%
  \noalign{%
    \vskip\belowdisplayshortskip
    \vtop{\hsize=\linewidth#1\par
      \expandafter}%
    \expandafter\prevdepth\the\prevdepth
  }%
}

\makeenvironment{Shaded}{\begin{snugshade}\small}{\end{snugshade}}
\makecommand{\AttributeTok}[1]{\textcolor[rgb]{0.42,0.37,0.12}{#1}}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Specifying models: The ODD protocol},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Specifying models: The ODD protocol}
\author{}
\date{\vspace{-2.5em}Reading for Class \#4: Monday, Sep 01, 2025}

\begin{document}
\maketitle

{
\setcounter{tocdepth}{2}
\tableofcontents
}
\subsection{Reading:}\label{reading}

\subsubsection{Required Reading
(everyone):}\label{required-reading-everyone}

\begin{itemize}
\tightlist
\item
  Agent-Based and Individual-Based Modeling, Ch. 3.
\end{itemize}

\subsubsection{Reading Notes:}\label{reading-notes}

Read carefully through the different design principles. Some of them
have meanings that are a bit different from what you might infer from
colloquial use.

For instance:

\begin{itemize}
\tightlist
\item
  \textbf{Adaptation} does not mean a persistent change in a turtle's
  behavior similar to the biological/Darwinian sense of adaptation in
  species. Rather, it means the way an agent changes its behavior in
  response to its \emph{immediate} conditions. Thus, adaptation in the
  ODD sense might include behaviors such as eating when you are hungry
  (\emph{eating} is an \textbf{adaptation} to \emph{hunger}), putting on
  warmer clothing when it's cold out (bundling up is an adaptation to
  cold), and running away from a predator.
\item
  The kind of persistent changes that arise over time from experience
  fall under the ODD design concept of \textbf{learning}: If there is
  more food near a river than on hills, turtles may \textbf{learn} to go
  to rivers when they are hungry.
\end{itemize}

You can download several useful documents related to the ODD protocol
from the class web site:

\begin{itemize}
\tightlist
\item
  The journal article, \href{/files/odd/Grimm_2010_ODD_update.pdf}{V.
  Grimm \emph{et al.} (2010). ``The ODD protocol: A review and first
  update'' \emph{Ecological Modeling} \textbf{221}, 2760--68.}.
  \url{https://ees4760.jgilligan.org/files/odd/Grimm_2010_ODD_update.pdf}
\item
  A Word document that provides
  \href{/files/odd/Grimm_2010_odd_template.docx}{a template for writing
  ODDs}:
  \url{https://ees4760.jgilligan.org/files/odd/Grimm_2010_odd_template.docx}
\item
  Lists of scientific publications using agent-based and
  individual-based models that either do or don't use the ODD protocol
  (this appeared as \href{/files/odd/Grimm_2010_appendix_1.pdf}{Appendix
  1} of the Grimm \emph{et al.} paper):

  \begin{itemize}
  \tightlist
  \item
    \url{https://ees4760.jgilligan.org/files/odd/Grimm_2010_appendix_1.pdf},
  \item
    \url{https://ees4760.jgilligan.org/files/odd/ch3_ex1_pubs_with_no_ODD.pdf},
  \item
    \url{https://ees4760.jgilligan.org/files/odd/ch3_ex2_pubs_with_ODD.pdf}.
  \end{itemize}
\end{itemize}

\end{document}
