---
title: "Reading Assignments Sheet"
author: "EES 3310/5310 Global Climate Change"
date: "Spring Semester, 2021."
output:
  pdf_document: 
    includes:
      in_header: assignment_header.sty
    keep_tex: yes
  html_document:
    df_print: paged
---

```{r setup, include=FALSE, message = FALSE, warning=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE)
```

# Disclaimer

This is a schedule of reading assignments through the entire term. 
I have worked hard to plan the semester, but I may need to deviate from this 
schedule, either because I decide that it's important to spend more time on 
some subjects, or because new developments in either climate science or 
climate policy require us to depart from my plans to discuss current events and 
breaking news.

The most up-to-date versions of the reading assignments will be posted on the
[course web site](https://ees3310.jgilligan.org): 
<https://ees3310.jgilligan.org>

# General instructions for reading assignments:


* Do the assigned reading _before_ you come to class on the date for 
  which it is assigned. If you have questions or find the ideas presented in 
  the readings confusing, I encourage you to ask questions in class.
* Questions in the ``Reading Notes'' sections of the assignments are for you to 
  think about to make sure you understand the material, but you do not have to 
  write up your answers or turn them in. On tests, you are responsible for all 
  the assigned readings, but topics I have highlighted in the reading notes are 
  particularly important.
* In addition to the questions I ask in the reading notes, look over the 
  ``study questions'' at the end of each chapter in _Understanding the Forecast_
  to check whether you understand the key facts and concepts from the chapter.

  Don't get confused between the **Study Questions** and the **Exercises** at 
  the end of the chapter: Study Questions are for your own use in reviewing 
  whether you understand the chapter, and the answers generally appear in the 
  text, so if you don't know the answer, look back at the chapter. Several of
  the laboratory assignments ask you to do the Exercises, which are more 
  challenging and ask you to apply the concepts from the chapter.

```{r initialize, include=FALSE}
library(semestr)

semester <- load_semester_db("../EES-3310-5310.sqlite3")

# output_to <- knitr::opts_knit$get("rmarkdown.pandoc.to")
# output_formats <- rmarkdown::all_output_formats(knitr::current_input())
# 
# if (output_to == "latex" || output_formats[[1]] == "pdf_document") {
#   .globals$text_mode <- "latex"
# } else {
#   .globals$text_mode <- "md"
# }

.globals <- new.env(parent = globalenv())

if (knitr::is_latex_output()) {
  assign("text_mode", "latex", envir = .globals)
} else {
  assign("text_mode", "md", envir = .globals)
}


```
```{r define_functions, include=FALSE}
assemble_schedule <- function(semester) {
  schedule <- init_schedule(semester)
  tmp <- schedule_strip_finals(schedule, semester)
  schedule <- tmp$schedule
  final_exams <- tmp$final_exams

  tmp <- schedule %>% schedule_add_homework(semester)
  schedule <- tmp$schedule

  tmp <- schedule_widen(schedule, final_exams, semester, TRUE)
  schedule <- tmp$schedule

  set_schedule_globals(schedule, semester)
  
  invisible(schedule)
}

format_section_heading <- function(title, class_date, class_number) {
  stringr::str_c("# ", class_date, ": ", stringr::str_trim(title))
}

make_rd_asgt_for_master <- function(id, schedule, semester) {
  reading <- semester$rd_items %>% dplyr::filter(cal_id == id) %>%
    # merge_dates(semester) %>%
    dplyr::left_join( dplyr::select(semester$class_topics, topic, rd_key),
               by = "rd_key")
  if (nrow(reading) >= 1) {
    rd_date <- unique(reading$date)
    assertthat::assert_that(length(rd_date) == 1,
                            msg = "A calendar ID should have a unique date (make_rd_for_master)")
    rd_topic <- unique(reading$topic)
    assertthat::assert_that(length(rd_topic) == 1,
                            msg = "A calendar ID should have a unique topic (make_rd_for_master)")
    class_num <- unique(reading$class_num)
    assertthat::assert_that(length(rd_date) == 1,
                            msg = "A calendar ID should have a unique class # (make_rd_for_master)")
    key <- unique(reading$rd_key)
    assertthat::assert_that(length(key) == 1,
                            msg = "A calendar ID should have a unique reading key # (make_rd_for_master)")
    if (is.na(rd_topic)) {
      cal_entry <- schedule %>% dplyr::filter(id_class == id)
      assertthat::assert_that(nrow(cal_entry) == 1, 
                              msg = "A calendar ID should have a unique schedule entry (make_rd_for_master)")
      rd_topic <- cal_entry$topic
    }
  } else {
    cal_entry <- schedule %>% dplyr::filter(id_class == id)
    assertthat::assert_that(nrow(cal_entry) == 1, 
                            msg = "A calendar ID should have a unique schedule entry (make_rd_for_master)")
    rd_date <- cal_entry$date
    rd_topic <- cal_entry$topic
    class_num <- cal_entry$class_num
    key <- cal_entry$key_class
  }

  this_reading_asgt <- make_reading_assignment(reading) %>% escape_dollar()
  
  notices <- semester$notices %>%
  dplyr::filter(cal_id == id, ! is.na(notice))

  this_notice <- make_notice(notices)

  section_heading <-  format_section_heading(
    title = rd_topic,
    class_date = format_class_day_date(rd_date, abbr_month = TRUE, abbr_wday = TRUE),
    class_number = class_num
  )
    
  # homework <- semester$hw_asgt %>% dplyr::filter(cal_id == !!cal_id) %>%
  #   # merge_dates(semester) %>%
  #   dplyr::left_join( dplyr::select(semester$hw_items, -hw_num, -cal_id),
  #                     by = "hw_key")
  # 
  # this_hw_asgt <- make_short_hw_assignment(homework) %>% escape_dollar()
  # 

  this_asgt <- cat_nl(section_heading, 
                      # this_hw_asgt,
                      this_reading_asgt,
                      this_notice,
                      start_par = TRUE, extra_lines = 1)
  this_asgt
}

make_holiday_entry_for_master <- function(id, schedule, semester) {
  holiday <- semester$holidays %>% dplyr::filter(cal_id == id) %>%
    # merge_dates(semester) %>%
    dplyr::left_join( dplyr::select(schedule, topic, holiday_key = key_holiday),
               by = "holiday_key") %>%
    dplyr::distinct()
  holiday_date <- unique(holiday$date)
  assertthat::assert_that(length(holiday_date) == 1,
                          msg = "A calendar ID should have a unique date (make_holiday_entry_for_master)")
  holiday_topic <- unique(holiday$topic)
  assertthat::assert_that(length(holiday_topic) == 1,
                          msg = "A calendar ID should have a unique topic (make_holiday_entry_for_master)")
  key <- unique(holiday$holiday_key)
  assertthat::assert_that(length(key) == 1,
                          msg = "A calendar ID should have a unique holiday key # (make_holiday_entry_for_master)")
  holiday_name <- unique(holiday$holiday_name)
  assertthat::assert_that(length(key) == 1,
                          msg = "A calendar ID should have a unique holiday name # (make_holiday_entry_for_master)")

  this_holiday_entry <- stringr::str_c(holiday_name, ", no class." )
  
  holiday_dates <- semester$holidays %>% 
    dplyr::filter(holiday_key == key) %$% date
  
  if (length(holiday_dates) == 1) {
    date_string <- format_class_day_date(holiday_date, abbr_month = TRUE, 
                                        abbr_wday = TRUE)
  } else {
    date_string <- format_day_date_range(holiday_dates, abbr_month = TRUE, 
                                        abbr_wday = TRUE)
  }
  
  section_heading <-  format_section_heading(
    title = holiday_topic,
    class_date = date_string,
    NA
  )
    
  
  # homework <- semester$hw_asgt %>% dplyr::filter(cal_id == !!cal_id) %>%
  #   # merge_dates(semester) %>%
  #   dplyr::left_join( dplyr::select(semester$hw_items, -hw_num, -cal_id),
  #                     by = "hw_key")
  # 
  # this_hw_asgt <- make_short_hw_assignment(homework) %>% escape_dollar()
  # 

  this_asgt <- cat_nl(section_heading, 
                      this_holiday_entry,
                      start_par = TRUE, extra_lines = 1)
  
  if (holiday_date > min(holiday_dates)) {
    this_asgt <- NA_character_
  }
  this_asgt
}


format_class_entry <- function(d, schedule, semester) {
  d <- lubridate::as_date(d)
  entry <- schedule %>% dplyr::filter(date == d)
  this_item <- NA_character_
  type <- NA_character_
  if (! is.na(entry$id_class)) {
    type <- "class"
    id <- entry$id_class
    this_item <- make_rd_asgt_for_master(entry$id_class, schedule, semester)
  } else if (! is.na(entry$id_holiday)) {
    type <- "holiday"
    id <- entry$id_holiday
    this_item <- make_holiday_entry_for_master(entry$id_holiday, schedule, semester)
  }
  dbg_checkpoint("g_date", d)
  if (! is.na(type) && ! is.na(this_item)) {
    key <- semester$calendar %>% dplyr::filter(cal_id == id) %$% cal_key
    context <- list(
      type = type,
      date = d,
      title = entry$topic,
      cal_key = key,
      cal_id = id,
      class_num = entry$class_num
    )
    this_item <- expand_codes(this_item, context, semester)
  }
  this_item
}

make_reading_schedule <- function(schedule, semester) {
  schedule <- schedule %>% dplyr::arrange(date)
  asgt_list <- character(0)
  for (d in schedule$date) {
    d <- lubridate::as_date(d)
    asgt_list <- c(asgt_list, 
                   format_class_entry(d, schedule, semester))
  }
  asgt_list <- asgt_list %>% purrr::discard(is.na) %>%
    cat_nl(start_par = TRUE)
}

```
```{r generate_assignments, echo=FALSE, include=TRUE, results="asis"}
schedule <- assemble_schedule(semester)  %>%
  dplyr::filter(! is.na(id_class) || ! is.na(id_holiday))

schedule <- make_reading_schedule(schedule, semester)
cat(schedule)
```

