---
title: "Reading Assignments"
author: "EES 4760/5760: Agent- and Individual-Based Computational Modeling"
date: "Fall, 2021"
output:
  pdf_document: 
    includes:
      in_header: assignment_header.sty
    keep_tex: yes
  html_document:
    df_print: paged
---
```{r setup, include=FALSE, message = FALSE, warning=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE)
```
```{r initialize, eval=TRUE, include=FALSE}
library(tidyverse)
library(semestr)

semester <- load_semester_db("../EES-4760.sqlite3")

# output_to <- knitr::opts_knit$get("rmarkdown.pandoc.to")
# output_formats <- rmarkdown::all_output_formats(knitr::current_input())
# 
# if (output_to == "latex" || output_formats[[1]] == "pdf_document") {
#   .globals$text_mode <- "latex"
# } else {
#   .globals$text_mode <- "md"
# }

.globals <- new.env(parent = globalenv())

if (knitr::is_latex_output()) {
  assign("text_mode", "latex", envir = .globals)
} else {
  assign("text_mode", "md", envir = .globals)
}


```

# General instructions for reading assignments

* Do the assigned reading \emph{before\/} you come to class on the date for 
  which it is assigned. If you have questions or find the ideas presented in 
  the readings confusing, I encourage you to ask questions in class.
* Check the list of errata at 
  <http://www.railsback-grimm-abm-book.com/downloads-errata-2nd-edition/>.
* Questions in the ``Reading Notes'' sections of the assignments are for you 
  to think about to make sure you understand the material, but you do not have 
  to write up your answers or turn them in.

  You are responsible for all the assigned readings, but topics I have 
  highlighted in the reading notes are particularly important.
* In addition to the questions I ask in the reading notes, look over the 
  "Conclusions" section at the end of each chapter in 
  _`r semester$text_codes$md['MED_RAILSBACK']`_
  to check whether you understand the key facts and concepts from the chapter.

## Disclaimer

This is a schedule of reading assignments through the entire term. 
I have worked hard to plan the semester, but I may need to deviate from this 
schedule, if I decide that it's important to spend more time on 
some subjects.

The most up-to-date versions of the reading assignments will be posted on the
"Schedule" page of the 
[course web site](https://ees4760.jgilligan.org/schedule): 
<https://ees4760.jgilligan.org/schedule>


```{r define_functions, eval=TRUE, include=FALSE, dependson="initialize"}
assemble_schedule <- function(semester) {
  schedule <- semestr:::init_schedule(semester)
  tmp <- semestr:::schedule_strip_finals(schedule, semester)
  schedule <- tmp$schedule
  final_exams <- tmp$final_exams

  tmp <- schedule %>% semestr:::schedule_add_homework(semester)
  schedule <- tmp$schedule

  tmp <- semestr:::schedule_widen(schedule, final_exams, semester, TRUE)
  schedule <- tmp$schedule

  semestr:::set_schedule_globals(schedule, semester)
  
  invisible(schedule)
}

format_section_heading <- function(title, class_date, class_number) {
  stringr::str_c("# ", class_date, ": ", stringr::str_trim(title))
}

make_rd_asgt_for_master <- function(id, schedule, semester) {
  reading <- semester$rd_items %>% dplyr::filter(cal_id == id) %>%
    # merge_dates(semester) %>%
    dplyr::left_join( dplyr::select(semester$class_topics, topic, rd_grp_key),
               by = "rd_grp_key")
  if (nrow(reading) >= 1) {
    rd_date <- unique(reading$date)
    assertthat::assert_that(length(rd_date) == 1,
                            msg = "A calendar ID should have a unique date (make_rd_for_master)")
    rd_topic <- unique(reading$topic)
    assertthat::assert_that(length(rd_topic) == 1,
                            msg = "A calendar ID should have a unique topic (make_rd_for_master)")
    class_num <- unique(reading$class_num)
    assertthat::assert_that(length(rd_date) == 1,
                            msg = "A calendar ID should have a unique class # (make_rd_for_master)")
    key <- unique(reading$rd_grp_key)
    assertthat::assert_that(length(key) == 1,
                            msg = "A calendar ID should have a unique reading key # (make_rd_for_master)")
    if (is.na(rd_topic)) {
      cal_entry <- schedule %>% dplyr::filter(id_class == id)
      assertthat::assert_that(nrow(cal_entry) == 1, 
                              msg = "A calendar ID should have a unique schedule entry (make_rd_for_master)")
      rd_topic <- cal_entry$topic
    }
  } else {
    cal_entry <- schedule %>% dplyr::filter(id_class == id)
    assertthat::assert_that(nrow(cal_entry) == 1, 
                            msg = "A calendar ID should have a unique schedule entry (make_rd_for_master)")
    rd_date <- cal_entry$date
    rd_topic <- cal_entry$topic
    class_num <- cal_entry$class_num
    key <- cal_entry$key_class
  }

  this_reading_asgt <- semestr:::make_reading_assignment(reading) %>% escape_dollar()
  
  if (! is.null(semester$notices)) {
    notices <- semester$notices %>%
      dplyr::filter(cal_id == id, ! is.na(notice))
    
    if (nrow(notices) > 0) {
      this_notice <- semestr:::make_notice(notices)
    }
  } else {
    this_notice <- NULL
  }
  
  section_heading <-  format_section_heading(
    title = rd_topic,
    class_date = format_class_day_date(rd_date, abbr_month = TRUE, abbr_wday = TRUE),
    class_number = class_num
  )
    
  # homework <- semester$hw_asgt %>% dplyr::filter(cal_id == !!cal_id) %>%
  #   # merge_dates(semester) %>%
  #   dplyr::left_join( dplyr::select(semester$hw_items, -hw_num, -cal_id),
  #                     by = "hw_key")
  # 
  # this_hw_asgt <- make_short_hw_assignment(homework) %>% escape_dollar()
  # 

  this_asgt <- semestr:::cat_nl(section_heading, 
                      # this_hw_asgt,
                      this_reading_asgt,
                      this_notice,
                      start_par = TRUE, extra_lines = 1)
  this_asgt
}

make_holiday_entry_for_master <- function(id, schedule, semester) {
  holiday <- semester$holidays %>% dplyr::filter(cal_id == id) %>%
    # merge_dates(semester) %>%
    dplyr::left_join( dplyr::select(schedule, topic, holiday_key = key_holiday),
               by = "holiday_key") %>%
    dplyr::distinct()
  holiday_date <- unique(holiday$date)
  assertthat::assert_that(length(holiday_date) == 1,
                          msg = "A calendar ID should have a unique date (make_holiday_entry_for_master)")
  holiday_topic <- unique(holiday$topic)
  assertthat::assert_that(length(holiday_topic) == 1,
                          msg = "A calendar ID should have a unique topic (make_holiday_entry_for_master)")
  key <- unique(holiday$holiday_key)
  assertthat::assert_that(length(key) == 1,
                          msg = "A calendar ID should have a unique holiday key # (make_holiday_entry_for_master)")
  holiday_name <- unique(holiday$holiday_name)
  assertthat::assert_that(length(key) == 1,
                          msg = "A calendar ID should have a unique holiday name # (make_holiday_entry_for_master)")

  this_holiday_entry <- stringr::str_c(holiday_name, ", no class." )
  
  holiday_dates <- semester$holidays %>% 
    dplyr::filter(holiday_key == key) %$% date
  
  if (length(holiday_dates) == 1) {
    date_string <- format_class_day_date(holiday_date, abbr_month = TRUE, 
                                        abbr_wday = TRUE)
  } else {
    date_range <- list(start = min(holiday_dates), stop = max(holiday_dates))
    date_string <- format_day_date_range(date_range, abbr_month = TRUE, 
                                        abbr_wday = TRUE)
  }
  
  section_heading <-  format_section_heading(
    title = holiday_topic,
    class_date = date_string,
    NA
  )
    
  
  # homework <- semester$hw_asgt %>% dplyr::filter(cal_id == !!cal_id) %>%
  #   # merge_dates(semester) %>%
  #   dplyr::left_join( dplyr::select(semester$hw_items, -hw_num, -cal_id),
  #                     by = "hw_key")
  # 
  # this_hw_asgt <- make_short_hw_assignment(homework) %>% escape_dollar()
  # 

  this_asgt <- semestr:::cat_nl(section_heading, 
                      this_holiday_entry,
                      start_par = TRUE, extra_lines = 1)
  
  if (holiday_date > min(holiday_dates)) {
    this_asgt <- NA_character_
  }
  this_asgt
}


format_class_entry <- function(d, schedule, semester) {
  d <- lubridate::as_date(d)
  entry <- schedule %>% dplyr::filter(date == d)
  this_item <- NA_character_
  type <- NA_character_
  if (! is.na(entry$id_class)) {
    type <- "class"
    id <- entry$id_class
    this_item <- make_rd_asgt_for_master(entry$id_class, schedule, semester)
  } else if (! is.na(entry$id_holiday)) {
    type <- "holiday"
    id <- entry$id_holiday
    this_item <- make_holiday_entry_for_master(entry$id_holiday, schedule, semester)
  }
  # dbg_checkpoint("g_date", d)
  if (! is.na(type) && ! is.na(this_item)) {
    key <- semester$calendar %>% dplyr::filter(cal_id == id) %$% cal_key
    context <- list(
      type = type,
      date = d,
      title = entry$topic,
      cal_key = key,
      cal_id = id,
      class_num = entry$class_num
    )
    this_item <- semestr:::expand_codes(this_item, context, semester)
  }
  this_item
}

make_reading_schedule <- function(schedule, semester) {
  schedule <- schedule %>% dplyr::arrange(date)
  asgt_list <- character(0)
  for (d in schedule$date) {
    d <- lubridate::as_date(d)
    asgt_list <- c(asgt_list, 
                   format_class_entry(d, schedule, semester))
  }
  asgt_list <- asgt_list %>% purrr::discard(is.na)
  asgt_list <- paste(asgt_list, collapse = "\n\n")
  invisible(asgt_list)
}

```
```{r generate_assignments, echo=FALSE, include=TRUE, results="asis", dependson="define_functions"}
schedule <- assemble_schedule(semester)  %>%
  dplyr::filter(! is.na(id_class) || ! is.na(id_holiday))

schedule <- make_reading_schedule(schedule, semester)
cat(schedule)
```

