---
title: "Homework Assignments"
author: "EES 4760/5760: Agent- and Individual-Based Computational Modeling"
date: "Fall, 2021"
output:
  pdf_document: 
    includes:
      in_header: assignment_header.sty
    keep_tex: yes
    keep_md: yes
  html_document:
    df_print: paged
---
```{r setup, include=FALSE, message = FALSE, warning=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE)
```
```{r initialize, eval=TRUE, include=FALSE}
library(tidyverse)
library(semestr)

semester <- load_semester_db("../EES-4760.sqlite3")

# output_to <- knitr::opts_knit$get("rmarkdown.pandoc.to")
# output_formats <- rmarkdown::all_output_formats(knitr::current_input())
# 
# if (output_to == "latex" || output_formats[[1]] == "pdf_document") {
#   .globals$text_mode <- "latex"
# } else {
#   .globals$text_mode <- "md"
# }

.globals <- new.env(parent = globalenv())

if (knitr::is_latex_output()) {
  assign("text_mode", "latex", envir = .globals)
} else {
  assign("text_mode", "md", envir = .globals)
}


```

# General instructions for homework assignments


* Homework must be turned in on Brightspace by 11:59 pm on the 
  due date unless the assignment gives a different time. 
  * Late homework will lose 5% for every day it is late, but will not lose
    more than 50%, no matter how late it is (you can always get up to half
    credit for completing late homework).
* When a homework assignment has more than one file, I encourage you to 
  use a program like 7Zip (Windows) or iZip or Keka (Mac) to make a Zip
  file containing all the different files for the assignment, and upload that 
  to Brightspace.

I encourage you to discuss homework assignments with your classmates. 
Some assignments will explicitly tell you to work in a team with a classmate. 
Even when the assignments do not specify working in teams, it is still fine to 
work together on homework assignments, but 
**unless the assignment tells you to work in a team, you must actually do all 
the work yourself**. 
This means that you can ask a classmate to explain how they solved a problem 
but you have to go through the steps independently and put the answer in your 
own words, not simply copy someone else's work.

**It is a violation of the honor code to turn in homework that someone 
else has done for you or which you copied from someone else.** 
If you are unsure about how the honor code applies to an assignment, please ask 
me.

## Homework Policy: Extra-Credit Options

Some homework problems are assigned to graduate students only.
Undergraduate students may choose to do any of these problems that interest 
them, and will receive for extra credit if they do.

I assign grades based on the required work, so your grade will not be affected 
by whether or not other students do extra credit problems (in other words, 
I never curve grades downward).

If you did poorly on some homework problems, or did not turn in some homework 
problems, you can make up some of that deficit by doing extra-credit
problems on subsequent assignments. Extra credit on homework only counts toward 
your homework grade for the semester and cannot compensate for missed work on 
other assignments, such as the team project or research project.


## Disclaimer

This is a schedule of homework assignments through the entire term. 
I have worked hard to plan the semester, but I may need to deviate from this 
schedule.

The most up-to-date versions of the homework assignments will be posted on the
"Schedule" page of the 
[course web site](https://ees4760.jgilligan.org/schedule): 
<https://ees4760.jgilligan.org/schedule>

```{r define_functions, eval=TRUE, include=FALSE, dependson="initialize"}
assemble_schedule <- function(semester) {
  schedule <- semestr:::init_schedule(semester)
  tmp <- semestr:::schedule_strip_finals(schedule, semester)
  schedule <- tmp$schedule

  tmp <- semestr:::schedule_add_homework(schedule, semester)
  schedule <- tmp$schedule

  tmp <- semestr:::schedule_widen(schedule, final_exams, semester, TRUE)
  schedule <- tmp$schedule

  semestr:::set_schedule_globals(schedule, semester)
  
  invisible(schedule)
}

format_section_heading <- function(hw_title, hw_date) {
  stringr::str_c("# ", hw_date, ": ", stringr::str_trim(hw_title))
}

make_hw_asgt_for_master <- function(id, schedule, semester) {
  homework <- semester$hw_asgt %>% dplyr::filter(cal_id == id)
  this_hw_asgt <- semestr:::make_hw_asgt_content(homework$hw_grp_key, 
                                                   semester) %>% 
    escape_dollar()
  
  section_heading <-  
    format_section_heading(hw_title = homework$hw_title,
                           hw_date = format_class_day_date(homework$date, 
                                                           abbr_month = TRUE, 
                                                           abbr_wday = TRUE)
    )

  this_asgt <- semestr:::cat_nl(section_heading, 
                                this_hw_asgt,
                                start_par = TRUE, extra_lines = 1)
  this_asgt
}

format_hw_entry <- function(d, schedule, semester) {
  d <- lubridate::as_date(d)
  entry <- schedule %>% dplyr::filter(date == d)
  this_item <- NA_character_
  type <- NA_character_
  if (! is.na(entry$id_hw)) {
    id <- entry$id_hw
    this_item <- make_hw_asgt_for_master(id, schedule, semester)
    hw_title <- semester$hw_asgt %>%
      dplyr::filter(cal_id == id) %>% pull(hw_title)
  } 
  # dbg_checkpoint("g_date", d)
  if (! is.na(this_item)) {
    key <- semester$calendar %>% dplyr::filter(cal_id == id) %>%
      pull(cal_key)
    context <- list(
      type = "homework",
      date = d,
      title = hw_title,
      cal_key = key,
      cal_id = id
    )
    this_item <- semestr:::expand_codes(this_item, context, semester)
  }
  this_item
}

make_homework_schedule <- function(schedule, semester) {
  schedule <- schedule %>% dplyr::arrange(date)
  asgt_list <- character(0)
  for (d in schedule$date) {
    d <- lubridate::as_date(d)
    asgt_list <- c(asgt_list, 
                   format_hw_entry(d, schedule, semester))
  }
  asgt_list <- asgt_list %>% purrr::discard(is.na)
  asgt_list <- paste(asgt_list, collapse = "\n\n")
  invisible(asgt_list)
}

```
```{r generate_assignments, echo=FALSE, include=TRUE, results="asis", dependson="define_functions"}
schedule <- assemble_schedule(semester)  %>%
  dplyr::filter(! is.na(id_class) || ! is.na(id_holiday))

schedule <- make_homework_schedule(schedule, semester)
cat(schedule)
```

